Predicting Counter-Strike Win Rates Using Player Statistics

### 2024-04-19 paper/paper.qmd
abstract and fixed spacing

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: 
  pdf: 
    fig-pos: 'H'
number-sections: true
bibliography: references.bib
---

@@

@fig-kdr shows the distribution of Kill/Death ratios. The median is slightly above the average of 1.0, 
indicating players in tournaments with significant prize pools are better than the average professional player, as expected. @fig-kd-1 and @fig-kd-2 shows the distributions of lifetime kills and deaths respectively.
```{r}
#| label: fig-kdr
#| fig-cap: Kill/Death ratios with median
#| echo: false
data |> 
  ggplot(aes(x = kill_death_ratio)) +
  geom_histogram(binwidth = 0.02) +
  labs(x = "Kill/Death Ratio", y = "Number of Players") +
  theme_classic() +
  geom_vline(aes(xintercept = median(kill_death_ratio)),
            color = "black",
            linetype = "dashed",
            linewidth = 1)
```

@@

```{r}
#| label: fig-adr
#| fig-cap: Distribution of ADR
#| echo: false
data |> 
  ggplot(aes(x = adr)) +
  geom_histogram(binwidth = 1) +
  labs(x = "ADR", y = "Number of Players") +
  theme_classic()
```

@fig-hs shows the distribution of headshot percentages. A professional *CS* team typically has 4 riflers and 1 AWPer[^1], where riflers aim for the head to do more damage while AWPers are more likely to aim for the chest since it is a larger area and they will kill in a single shot regardless. This could explain why the players are distributed with a left skew.
[^1]: An AWP is a powerful but expensive sniper rifle, capable of killing in one shot to the chest
```{r}
#| label: fig-hs
#| fig-cap: Distribution of Headshot Percentage
#| echo: false
data |> 
  ggplot(aes(x = hs)) +
  geom_histogram(binwidth = 1) +
  labs(x = "Headshot %", y = "Number of Players") +
  theme_classic()
```
@fig-rounds shows the total number of rounds played.
```{r}
#| label: fig-rounds
#| fig-cap: Number of rounds played
#| echo: false
data |> 
  ggplot(aes(x = rounds)) +
  geom_histogram(binwidth = 1000) +
  labs(x = "Rounds Played", y = "Number of Players") +
  theme_classic()
```

@@

# Results {#sec-results}
@fig-kdr-wr shows the relationship between K/D and map win rates with a line provided by the linear regression model.
The overall relationships in this figure and following figures are difficult to see using only data points due to the number and distribution.
However, the regression line suggests there is a positive correlation between K/D and win rate, as expected when constructing the model.

```{r}
#| label: fig-kdr-wr
#| fig-cap: Win rate by kill/death ratio with regression line
#| echo: false
plot_predictions(winrates, condition = "kill_death_ratio") +
  geom_point(data = data, aes(x = kill_death_ratio, y = win_percent), alpha = 0.5) +
  labs(x = "Kill/Death Ratio", y = "Win Rate") +
  theme_classic()
```

@fig-hs-wr shows the relationship between HS% and map win rates with a line provided by the linear regression model.
The regression line suggests there is a positive correlation between HS% and win rate, as expected when constructing the model.
```{r}
#| label: fig-hs-wr
#| fig-cap: Win rate by headshot percentage with regression line
#| echo: false
plot_predictions(winrates, condition = "hs") +
  geom_point(data = data, aes(x = hs, y = win_percent), alpha = 0.5) +
  labs(x = "Headshot Percentage", y = "Win Rate") +
  theme_classic()
```
@fig-adr-wr shows the relationship between ADR and map win rates with a line provided by the linear regression model.
Interestingly, the regression line suggests there is a negative correlation between ADR and win rate.
This appears to be counterintuitive since more damage is generally good. An explanation for this could be 
overaggressive players deal more damage but leave their teams in a worse position, leading to lower win rates. 
Another explanation could be "baiting", where a player waits for their team members to distract enemies in order to have easier kills,
allowing them to stay alive, inflating their personal stats but reducing the chance for their team to win.
```{r}
#| label: fig-adr-wr
#| fig-cap: Win rate by ADR with regression line
#| echo: false
plot_predictions(winrates, condition = "adr") +
  geom_point(data = data, aes(x = adr, y = win_percent), alpha = 0.5) +
  labs(x = "Average Damage per Round", y = "Win Rate") +
  theme_classic()
```
@fig-rounds-wr shows the relationship between total number of rounds played and map win rates with a line provided by the linear regression model.
The regression line suggests there is a positive relationship between total rounds played and win rate, as expected when constructing the model.
```{r}
#| label: fig-rounds-wr
#| fig-cap: Win rate by total rounds played with regression line
#| echo: false
plot_predictions(winrates, condition = "rounds") +
  geom_point(data = data, aes(x = rounds, y = win_percent), alpha = 0.5) +
  labs(x = "Rounds Played", y = "Win Rate") +
  theme_classic()
```
@tbl-model shows the model coefficients. Note this is not standardized, so the coefficient for K/D is large since a $\pm$ 0.1 change in K/D is a large difference whereas the coefficient for number of rounds is tiny since a change of $\pm$ 1 round played is practically nothing.

@@

modelsummary(
  list("Tournament Map Win Rate" = winrates),
  statistic = "mad",
  fmt = 6
  )
```

# Discussion{#sec-discussion}

>>>>

---
title: "Predicting Counter-Strike Win Rates Using Player Statistics"
author: "Shuyang Qiu"
thanks: "Code and data are available at: https://github.com/shuyangqiu/cs_stats."
date: today
date-format: long
abstract: "Counter-Strike player statistics were used to create a model to predict their team's win rate. Predictors used include player performance metrics such as
  average damage per round and overall statistics such as total number of rounds played. We found some average damage per round is negatively correlated with win rates
  while other metrics were positively correlated. This model could be improved on to allow bookmakers to create more efficient markets, or allow coaches to make better roster decisions."
format: 
  pdf: 
    fig-pos: 'H'
number-sections: true
bibliography: references.bib
---

@@

\newpage
@fig-kdr shows the distribution of Kill/Death ratios. The median is slightly above the average of 1.0, 
indicating players in tournaments with significant prize pools are better than the average professional player, as expected. @fig-kd-1 and @fig-kd-2 shows the distributions of lifetime kills and deaths respectively.
```{r}
#| label: fig-kdr
#| fig-cap: Kill/Death ratios with median
#| echo: false
data |> 
  ggplot(aes(x = kill_death_ratio)) +
  geom_histogram(binwidth = 0.02) +
  labs(x = "Kill/Death Ratio", y = "Number of Players") +
  theme_classic() +
  geom_vline(aes(xintercept = median(kill_death_ratio)),
            color = "black",
            linetype = "dashed",
            linewidth = 1)
```

@@

```{r, out.height = "75%", out.width = "75%"}
#| label: fig-adr
#| fig-cap: Distribution of ADR
#| echo: false
data |> 
  ggplot(aes(x = adr)) +
  geom_histogram(binwidth = 1) +
  labs(x = "ADR", y = "Number of Players") +
  theme_classic()
```

@fig-hs shows the distribution of headshot percentages. A professional *CS* team typically has 4 riflers and 1 AWPer[^1], where riflers aim for the head to do more damage while AWPers are more likely to aim for the chest since it is a larger area and they will kill in a single shot regardless. This could explain why the players are distributed with a left skew.

[^1]: An AWP is a powerful but expensive sniper rifle, capable of killing in one shot to the chest
```{r, out.height = "75%", out.width = "75%"}
#| label: fig-hs
#| fig-cap: Distribution of Headshot Percentage
#| echo: false
data |> 
  ggplot(aes(x = hs)) +
  geom_histogram(binwidth = 1) +
  labs(x = "Headshot %", y = "Number of Players") +
  theme_classic()
```
@fig-rounds shows the total number of rounds played.
```{r, out.height = "75%", out.width = "75%"}
#| label: fig-rounds
#| fig-cap: Number of rounds played
#| echo: false
data |> 
  ggplot(aes(x = rounds)) +
  geom_histogram(binwidth = 1000) +
  labs(x = "Rounds Played", y = "Number of Players") +
  theme_classic()
```

@@

\newpage
# Results {#sec-results}
@fig-kdr-wr shows the relationship between K/D and map win rates with a line provided by the linear regression model.
The overall relationships in this figure and following figures are difficult to see using only data points due to the number and distribution.
However, the regression line suggests there is a positive correlation between K/D and win rate, as expected when constructing the model.

```{r}
#| label: fig-kdr-wr
#| fig-cap: Win rate by kill/death ratio with regression line
#| echo: false
plot_predictions(winrates, condition = "kill_death_ratio") +
  geom_point(data = data, aes(x = kill_death_ratio, y = win_percent), alpha = 0.5) +
  labs(x = "Kill/Death Ratio", y = "Win Rate") +
  theme_classic()
```
\newpage
@fig-hs-wr shows the relationship between HS% and map win rates with a line provided by the linear regression model.
The regression line suggests there is a positive correlation between HS% and win rate, as expected when constructing the model.
```{r}
#| label: fig-hs-wr
#| fig-cap: Win rate by headshot percentage with regression line
#| echo: false
plot_predictions(winrates, condition = "hs") +
  geom_point(data = data, aes(x = hs, y = win_percent), alpha = 0.5) +
  labs(x = "Headshot Percentage", y = "Win Rate") +
  theme_classic()
```
\newpage
@fig-adr-wr shows the relationship between ADR and map win rates with a line provided by the linear regression model.
Interestingly, the regression line suggests there is a negative correlation between ADR and win rate.
This appears to be counterintuitive since more damage is generally good. An explanation for this could be 
overaggressive players deal more damage but leave their teams in a worse position, leading to lower win rates. 
Another explanation could be "baiting", where a player waits for their team members to distract enemies in order to have easier kills,
allowing them to stay alive, inflating their personal stats but reducing the chance for their team to win.
```{r}
#| label: fig-adr-wr
#| fig-cap: Win rate by ADR with regression line
#| echo: false
plot_predictions(winrates, condition = "adr") +
  geom_point(data = data, aes(x = adr, y = win_percent), alpha = 0.5) +
  labs(x = "Average Damage per Round", y = "Win Rate") +
  theme_classic()
```
\newpage
@fig-rounds-wr shows the relationship between total number of rounds played and map win rates with a line provided by the linear regression model.
The regression line suggests there is a positive relationship between total rounds played and win rate, as expected when constructing the model.
```{r}
#| label: fig-rounds-wr
#| fig-cap: Win rate by total rounds played with regression line
#| echo: false
plot_predictions(winrates, condition = "rounds") +
  geom_point(data = data, aes(x = rounds, y = win_percent), alpha = 0.5) +
  labs(x = "Rounds Played", y = "Win Rate") +
  theme_classic()
```
@tbl-model shows the model coefficients. Note this is not standardized, so the coefficient for K/D is large since a $\pm$ 0.1 change in K/D is a large difference whereas the coefficient for number of rounds is tiny since a change of $\pm$ 1 round played is practically nothing.

@@

modelsummary(
  list("Tournament Map Win Rate" = winrates),
  statistic = "mad",
  fmt = 6
  )
```
\newpage
\newpage
# Discussion{#sec-discussion}
